% IMPORTANT: add or remove (comment out) the boolean '\solutiontrue' below to
% create the solution document or the exercise document respectively.
% First we create the switch to make either the exercises or the solutions
\newif\ifsolution\solutionfalse
% To create the solution uncomment '\solutiontrue'
\solutiontrue

\documentclass[a4paper,11pt]{article}

\title{System Security,
\ifsolution Solution \else \fi
Lab Report}

\author{List group members here.}

\usepackage[T1]{fontenc}
\usepackage{ae, aecompl}
\usepackage{a4wide}
\usepackage{boxedminipage}
\usepackage{url}
\usepackage{graphicx}
\usepackage{enumerate}

% Some useful commands and environments
\usepackage{framed}
\newenvironment{solution}%
{\par{\noindent\small\textit{Solution:}}\vspace{-12pt}\begin{framed}}%
{\end{framed}\par}

\begin{document}
\maketitle

\section{General Hints for the reports}

To help you structuring the report on the labs, we provide you with some
questions to answer in the report and some \textbf{hints} you might need to
consider. Please, note that it is not an exhaustive list of questions.
Therefore, we advise you to include more information from your hacking
experience in the report.

\section{Lab Session: Side-channel attack}
In the lab session, the goal is to find the secret key used in a RSA
operation by analyzing the power consumption.
\begin{enumerate}[(a)]
\item Explain the setup to measure the power consumption of the sensor node
\ifsolution\begin{solution}
WRITE HERE.
\end{solution}\fi

\item To which values did you set the oscilloscope (e.g., horizontal
resolution) in order to observe the power trace? How did you find
these values?
\ifsolution\begin{solution}
WRITE HERE.
\end{solution}\fi

\item What technique did you use to improve the measurement noise in
the power trace?
\ifsolution\begin{solution}
WRITE HERE.
\end{solution}\fi

\item What is the key that you could see? Explain how you are able to
see the key and what leads to differences in the power consumption.
\ifsolution\begin{solution}
WRITE HERE.
\end{solution}\fi

\end{enumerate}


\end{document}
